\chapter*{Conclusion}
\addcontentsline{toc}{chapter}{Conclusion}

During the work on this thesis, we first analyzed the connection search problem and its existing solutions. We then used this information to set the requirements for our new app. To reach our goal of creating a more personalizable app, we added new requirements to solve some of the shortcomings of other existing app. We then analyzed what data is available to help us build the application and how to use it.

After this, we designed our application. We analyzed the algorithms that could be used to solve the problem and chose the most appropriate one. We went over programming languages that could be used for the project and chose the most suitable one. We have also chosen what high-level our application was going to have.

We implemented the application, which consists of a server and a client part. We documented the source code in detail, both inside the code-base using documentation comments and in this document, describing the structure of the code. We went over different decisions faced during the implementation and the solutions chosen. We also summarized information on the usage of external third-party software within our code. Lastly, we also explained in detail how to use the application, including documenting the installation process and usage of both the client and server apps. We also documented the usage of the server-side API.

Finally, we have tested our application in multiple ways. This includes automatized unit tests testing different components of the server application, tests checking the integrity of results returned by the server-side application, performance tests to find out the response times of the server application and usability tests to check that the client application provides the functionality it should.

We have achieved our goal of creating an application that offers more personalization and customization options than the existing solutions. We have also successfully incorporated bikesharing into our searches, being only the second application in Prague to do so. There are, however, still many opportunities for future expansion. These include implementing the client app for the iOS operating system as well, implementing useful features that other apps have, such as displaying the resulting connections on a map, displaying the exact stops every trip takes or providing information on the cost of tickets necessary to perform the connections. Thus, we might take this work and expand it in the future to provide a more robust, all-round solution than the scope of this work allowed us to.
